Learning Action Models from State Observations

### 2018-01-30 doc/ijcai-18/observations-learning.tex
Add files via upload

For the training samples, we adopt a middle ground between unstructured inputs and plan traces, wherein only state observations are required. \textbf{Learning from state observations} is a relevant advancement as in many applications, the actual actions executed by the observed agent are not available but instead, the resulting states can be observed. Learning action models from state observations broadens the range of application to external observers and facilitates the representation of imperfect observability (as shown in plan recognition \cite{SohrabiRU16}) as well as learning from unstructured data (like state images \cite{AsaiF18}).

@@

Unlike the error function defined by {\sc ARMS}~\cite{yang2007learning} the {\em observation edit distance} assess, with a single expression, the flaws in the preconditions and effects of a given learned model. This fact enables the recognition of \strips\ action models. The idea, taken from {\em plan recognition as planning}~\cite{ramirez2009plan}, is to map distances into likelihoods. The {\em observation edit distance} could be mapped into a likelihood with the following expression $P(\mathcal{O}|\mathcal{M})=1-\frac{\delta(\mathcal{M},\mathcal{O})}{\delta(\mathcal{M},*)}$.

@@

To compute the {\em observation edit distance}, our interest is not in $\mathcal{M}'$ but in the number of required {\em edit operations} for that $\mathcal{M}'$ is validated in the given observations, e.g. $\delta(\mathcal{M},\mathcal{O})=2$ for the example in Figure~\ref{fig:plan}. The {\em observation edit distance} is exactly computed if $P_{\Lambda}'$ is optimally solved (according to the number of edit actions); is approximated if $P_{\Lambda}'$ is solved with a satisfying planner (our case); and is a less accurate estimate (but faster to be computed) if the solved task is a relaxation of $P_{\Lambda}'$~\cite{bonet2001planning}.
\begin{figure}
{\tt\small
00 : (insert\_add\_handempty\_stack)\\
01 : (insert\_add\_clear\_stack\_var1)\\
02 : (apply\_unstack block2 block1)\\
03 : (validate\_1)\\
04 : (apply\_putdown block2)\\
05 : (validate\_2)\\
06 : (apply\_pickup block1)\\
07 : (validate\_3)\\
08 : (apply\_stack block1 block2)\\
09 : (validate\_4)\\
}
 \caption{\small Plan for editing a given {\em blockswold} schema and validating it at the state observations shown in Figure~\ref{fig:lexample}.}
\label{fig:plan}
\end{figure}

>>>>

For the training samples, we adopt a middle ground between unstructured inputs and plan traces, wherein only state observations are required. \textbf{Learning from state observations} is a relevant advancement as, in many applications, the actual actions executed by the observed agent are not available but, instead, the resulting states can be observed. Learning action models from state observations broadens the range of application to external observers and facilitates the representation of imperfect observability (as shown in plan recognition \cite{SohrabiRU16}) as well as learning from unstructured data (like state images \cite{AsaiF18}).

@@

Unlike the error function defined by {\sc ARMS}~\cite{yang2007learning}, the {\em observation edit distance} assesses, with a single expression, the flaws in the preconditions and effects of a given learned model. This fact enables the recognition of \strips\ action models. The idea, taken from {\em plan recognition as planning}~\cite{ramirez2009plan}, is to map distances into likelihoods. The {\em observation edit distance} could be mapped into a likelihood with the following expression $P(\mathcal{O}|\mathcal{M})=1-\frac{\delta(\mathcal{M},\mathcal{O})}{\delta(\mathcal{M},*)}$.

@@

Our interest when computing the {\em observation edit distance} is not in $\mathcal{M}'$ but in the number of required {\em edit operations} for that $\mathcal{M}'$ is validated in the given observations, e.g. $\delta(\mathcal{M},\mathcal{O})=2$ for the example in Figure~\ref{fig:plan}. The {\em observation edit distance} is exactly computed if $P_{\Lambda}'$ is optimally solved (according to the number of edit actions); is approximated if $P_{\Lambda}'$ is solved with a satisfying planner (our case); and is a less accurate estimate (but faster to be computed) if the solved task is a relaxation of $P_{\Lambda}'$~\cite{bonet2001planning}.

\begin{figure}
{\tt\small
00 : (insert\_add\_handempty\_stack)\\
01 : (insert\_add\_clear\_stack\_var1)\\
02 : (apply\_unstack block2 block1)\\
03 : (validate\_1)\\
04 : (apply\_putdown block2)\\
05 : (validate\_2)\\
06 : (apply\_pickup block1)\\
07 : (validate\_3)\\
08 : (apply\_stack block1 block2)\\
09 : (validate\_4)\\
}
 \caption{\small Plan for editing a given {\em blockswold} schema and validating it at the state observations shown in Figure~\ref{fig:lexample}.}
\label{fig:plan}
\end{figure}
